% Define the document as a subfile and reference its root.
\documentclass[../main.tex]{subfiles}

% Begin the document.
\begin{document}
\begin{preliminary}

    \title{A Graphical Playground for Haskell}
    \author{Oliver Jones}
    \course{Computer Science}
    \project{4th Year Project Report}
    \date{\today}

    \abstract{
        Functional programming is a fundamental paradigm in computer science education,
        yet it often presents a steep learning curve for students unfamiliar with its concepts.
        This project introduces \emph{A Graphical Playground for Haskell}, an
            interactive online environment, facilitating the learning and exploration of
            Haskell through a custom-built graphics library, available at
            \url{https://haskell-playground.co.uk}.
        The platform allows users to write, execute and share Haskell programs, to
            generate images and animations which can be viewed directly in the browser.

        This dissertation details the design and implementation of both the graphics
            library and the website.
        The system is structured to securely execute Haskell code on a server, and
            stream the results back to the client.
        Key challenges addressed include ensuring safe code execution and designing an
            intuitive user interface.

        User testing was conducted to evaluate the usability and effectiveness of the
            tool.
        The results indicate that the platform can successfully facilitate the learning
            of Haskell, with users reporting a positive experience using the system.
        The project demonstrates the potential of interactive environments in enhancing
            the learning of functional programming languages.
    }

    \maketitle

    \newenvironment{ethics}{
        \begin{frontenv}{Research Ethics Approval}{\LARGE}} {
        \end{frontenv}
        \newpage}

    \begin{ethics}
        \raggedright
        This project obtained approval from the Informatics Research Ethics committee.
        \\
        Ethics application number: 8978
        \\
        Date when approval was obtained: 2025-01-08
        \\

        The participants' information sheet and a consent form are included in
            Appendix~\ref{app:consent}.
        \\

        \standarddeclaration
    \end{ethics}

    \begin{acknowledgements}
        I would like to extend my most sincere thanks to my supervisor, Prof.~Philip
            Wadler, for his unwavering guidance and support throughout the project.
        I would also like to extend my gratitude to all those who tested the software
            and provided feedback, particularly those who completed the user testing
            survey.
    \end{acknowledgements}

    \tableofcontents
\end{preliminary}
\end{document}
